---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
Clean & load Libraries
```{r}
rm(list=ls())
library(dplyr)
library(lubridate)
library(ggplot2)
```
Loading
```{r}
setwd("~/Desktop")
activity <- read.csv("activity.csv", header=TRUE, na.strings = "NA")
```
reasign date class
```{r}
activity$date <- ymd(activity$date)
```
Remove NA
```{r}
activityclean <- na.omit(activity)
```
## What is mean total number of steps taken per day?
Summarize data
```{r}
activitysum <- summarize(group_by(activityclean,date),daily.steps=sum(steps))
activitysum 
```
(Gives the total number of steps taken per day)

Plot histogram
```{r 1.Histogram}
histplot = qplot(activitysum$daily.steps, binwidth = 1000, main = "Total number of steps taken each day", xlab= "Daily Steps", ylab = "Frequency", col=I("red"), alpha=I(.2), fill=I("blue"))
histplot
```

Calculating Mean & Median (it is necessary to coerce to an integer)
```{r}
stepsmean = as.integer (mean(activitysum$daily.steps)) 
stepsmean 
```
Total number of steps taken per day is 10766
```{r}
stepsmedian = as.integer (median(activitysum$daily.steps))
stepsmedian 
```
Total number of steps taken per day is 10765

## What is the average daily activity pattern?
Summarize data
```{r}
aveday <-summarize(group_by(activityclean,interval),mean.steps=mean(steps))
```
Plot average number of steps by 5-minute interval
```{r 1.Timeseries}
seriesplot <- ggplot(data = aveday, aes(x = interval, y = mean.steps)) + geom_line(color="red")+ labs(title="Average Number of Steps Taken vs 5-min Interval", y="Average Number of Steps", x="5-min Interval")
seriesplot
```

Calculating which 5-minute interval,contains the maximum number of steps
```{r}
maxposition <- which.max(aveday$mean.steps)
peak <- aveday$interval[maxposition]
peak 
```
Maximum number of steps is 835

## Imputing missing values
Calculating missing values
```{r}
sum(is.na(activity)) 
```
Total number of missing values in the origina dataset is 2304

Devise a strategy for filling in all of the missing values in the dataset:
NA's will be substituted by the steps' mean value (without NA's)
```{r}
filled.activity <- activity
filled.activity$steps[is.na(filled.activity$steps)] <- mean(filled.activity$steps,na.rm=TRUE)
```
Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
colSums(is.na(filled.activity)) # To check there is no "NA's"
```
To check there is no "NA's"

Creating New Histogram:

Summarize data
```{r}
filled.activity2 <- summarize(group_by(filled.activity,date),daily.steps=sum(steps))
```
Calculating the Mean
```{r}
mean.filled   <- as.integer(mean(filled.activity2$daily.steps))
mean.filled 
```
Mean = 10766

Calculating the Median
```{r}
median.filled   <- as.integer(median(filled.activity2$daily.steps))
median.filled 
```
Median = 10766

Plotting
```{r 2.Histogram}
histplot2 = qplot(filled.activity2$daily.steps, binwidth = 1000, main = "Total number of steps taken each day", xlab= "Daily Steps", ylab = "Frequency", col=I("red"), alpha=I(.2), fill=I("blue"))
histplot2
```

After NA's substitution, the Mean and Median become the same. 
Filling with mean values make the histogram more normally distributed.

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
```{r}
filled.activity$day <- ifelse(weekdays(filled.activity$date) %in% c("Saturday","Sunday"), "weekday", "weekend")
```
Preparing data for ggplot (Summarizing)
```{r}
filled.activity2 <- summarize(group_by(filled.activity,interval,day),steps.mean=mean(steps))
```
Plotting
```{r 2.Timeseries}
weekday.interval <- ggplot (filled.activity2, aes(x=interval, y=steps.mean, color=day)) + facet_grid(day~.) + geom_line() + labs (title="Average number of steps taken in 5-min interval", y="Average number of steps", x="5-min interval")
weekday.interval
```


